---
title: "ESM 260 Problem Set 1"
author: "E Griffiths"
date: "2024-01-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(broom)

adult_fish <- read_csv("adult_fish.csv") %>%
  drop_na()
as.numeric("Density")
```

Question 1: Make a histogram that displays the mean density of the organism (y-axis) at Rarotonga and Aitutaki for 2009 and for 2024. Add the 95% confindence intervals for each species.

a) Surgeonfish

```{r}
adult_surgeonfish <- adult_fish %>%
  filter(Fish == "Surgeonfish")

adult_surgeonfish_plot <- ggplot(adult_surgeonfish, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Adult Surgeonfish") +
  guides(fill = guide_legend(title = "Island")) +
  geom_errorbar( aes(ymin = Density - 32, ymax = Density + 32), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
  theme_minimal() 
adult_surgeonfish_plot

ggsave("adult_surgeonfish_plot.jpeg", width = 7, height = 4, units = "in")
```

b) Coral Trout

```{r}
adult_coraltrout <- adult_fish %>%
  filter(Fish == "Coral Trout")

adult_coraltrout_plot <- ggplot(adult_coraltrout, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
 geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Adult Coral Trout") +
  guides(fill = guide_legend(title = "Island")) +
 geom_errorbar( aes(ymin = Density - 21, ymax = Density + 21), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
 theme_minimal() 
adult_coraltrout_plot

ggsave("adult_coraltrout_plot.jpeg", width = 7, height = 4, units = "in")
```

c) Spotted Damselfish

```{r}
adult_spotteddamsel <- adult_fish %>%
  filter(Fish == "Spotted Damselfish")

adult_spotteddamsel_plot <- ggplot(adult_spotteddamsel, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
 geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Adult Spotted Damselfish") +
  guides(fill = guide_legend(title = "Island")) +
 geom_errorbar( aes(ymin = Density - 47, ymax = Density + 47), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
 theme_minimal() 
adult_spotteddamsel_plot

ggsave("adult_spotteddamsel_plot.jpeg", width = 7, height = 4, units = "in")
```
d)

```{r}
adult_yellowdamsel <- adult_fish %>%
  filter(Fish == "Yellow Damselfish")

adult_yellowdamsel_plot <- ggplot(adult_yellowdamsel, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
 geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Adult Yellow Damselfish") +
  guides(fill = guide_legend(title = "Island")) +
 geom_errorbar( aes(ymin = Density - 15, ymax = Density + 15), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
 theme_minimal() 
adult_yellowdamsel_plot

ggsave("adult_yellowdamsel_plot.jpeg", width = 7, height = 4, units = "in")
```

Question 2: Do the same as in Question 1 for the recruits (young) of each fish species.  

```{r}
young_fish <- read_csv("juve_fish.csv") %>%
drop_na()
as.numeric("Density")
```

a) Surgeonfish

```{r}
young_surgeonfish <- young_fish %>%
  filter(Fish == "Surgeonfish")

young_surgeonfish_plot <- ggplot(young_surgeonfish, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Young Surgeonfish") +
  guides(fill = guide_legend(title = "Island")) +
  geom_errorbar( aes(ymin = Density - 10.0, ymax = Density + 10.0), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
  theme_minimal() 
young_surgeonfish_plot

ggsave("young_surgeonfish_plot.jpeg", width = 7, height = 4, units = "in")
```

b) Coral Trout

```{r}
young_coraltrout <- young_fish %>%
  filter(Fish == "Coral Trout")

young_coraltrout_plot <- ggplot(young_coraltrout, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
 geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Young Coral Trout") +
  guides(fill = guide_legend(title = "Island")) +
 geom_errorbar( aes(ymin = Density - 2.5, ymax = Density + 2.5), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
 theme_minimal() 
young_coraltrout_plot

ggsave("young_coraltrout_plot.jpeg", width = 7, height = 4, units = "in")
```

c) Spotted Damselfish

```{r}
young_spotteddamsel <- young_fish %>%
  filter(Fish == "Spotted Damselfish")

young_spotteddamsel_plot <- ggplot(young_spotteddamsel, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
 geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Young Spotted Damselfish") +
  guides(fill = guide_legend(title = "Island")) +
 geom_errorbar( aes(ymin = Density - 5.0, ymax = Density + 5.0), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
 theme_minimal() 
young_spotteddamsel_plot

ggsave("young_spotteddamsel_plot.jpeg", width = 7, height = 4, units = "in")
```

d) Yellow Damselfish

```{r}
young_yellowdamsel <- young_fish %>%
  filter(Fish == "Yellow Damselfish")

young_yellowdamsel_plot <- ggplot(young_yellowdamsel, aes(x = as.factor(Year), y = Density, fill = as.factor(Island))) +
 geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Year", y = "Density (per 100m2)", title = "Young Yellow Damselfish") +
  guides(fill = guide_legend(title = "Island")) +
 geom_errorbar( aes(ymin = Density - 4.0, ymax = Density + 4.0), width=0.4, colour="black", width = 0.08, position = position_dodge(0.8)) +
 theme_minimal() 
young_yellowdamsel_plot

ggsave("young_yellowdamsel_plot.jpeg", width = 7, height = 4, units = "in")
```









